---
title: "Visualization pt 1"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
```

# Load the weather data 

```{r}

weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") %>% 
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>% 
  select(name, id, everything())

weather_df
```

## Scatterplots

Create scatterplot 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

New approach, same scatterplot: 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Save and edit a plot object (a frame to save the plot):

```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))
  
weather_plot + geom_point()
```

## Advanced scatterplot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE)  
```

What about the placement of 'aes'...

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) + 
  geom_smooth()
```

...its generally best practice to put the color specification with the plot rather than geom_point bc in the latter it is only applied to the actual creation of the scatterplot graph 

Lets facet some things 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .2) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

Lets combine some elements and try a new plot 

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) + 
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

## Some small notes

How many geoms need to exist? --you can have as many as you want

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name))
```

You can use a 'neat' geom 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex() +
  geom_density_2d()
```

## Univariate plots 

Histograms

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram() 
```

Adding color to histograms

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

Comparing distributions across a certain level - best geom: 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .3)
```

What about box plots?

```{r}
weather_df %>% 
  ggplot(aes(y = tmin, x = name)) +
  geom_boxplot()
```

~Trendy plots~

```{r}
weather_df %>% 
  ggplot(aes(y = tmin, x = name, fill = name)) +
  geom_violin() +
  stat_summary()
```
--> violin plots can be more informative than a boxplot at times

Ridge plots -- good for when you have a categorical predictor with lots of variables

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```

## Save and embed 

Save a scatterplot

```{r}
weather_plot =
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)

ggsave("weather_plot.png", weather_plot)
```

What about embedding?

```{r}
weather_plot
```

Embed at a different size
```{r, fig.width=12, fig.height=6}
weather_plot
```
-- Need to be aware of what size you are embedding things at to ensure that things look okay in the output 