---
title: "Visualization pt 2"
output: github_document
---

```{r}
library(tidyverse)
library(patchwork)
```

## Load the weather data 

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2023-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

```

## Using labs() to label this plot
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum Daily Temperature (C)",
    y = "Maximum Daily Temperature (C)",
    caption = "Data from rnoaa package; temps in 2017"
  )
  
```

## Scales 

Start w the same plot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum Daily Temperature (C)",
    y = "Maximum Daily Temperature (C)",
    caption = "Data from rnoaa package; temps in 2017"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C","0 C", "15 C")
  ) +
  scale_y_continuous(
    position = "right" 
  )
```

## Look at color scales 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum Daily Temperature (C)",
    y = "Maximum Daily Temperature (C)",
    caption = "Data from rnoaa package; temps in 2017"
  ) +
 viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE)
```

## Themes 

Shift the legend: 
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum Daily Temperature (C)",
    y = "Maximum Daily Temperature (C)",
    caption = "Data from rnoaa package; temps in 2017"
  ) +
 viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
  theme(legend.position = "bottom")
```

Change overall theme (theme minimal is jeffs fav)
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum Daily Temperature (C)",
    y = "Maximum Daily Temperature (C)",
    caption = "Data from rnoaa package; temps in 2017"
  ) +
 viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Setting options

```{r, eval=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuious.color = "viridis",
  ggplot2.continuious.fill = "viridis"
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Data arguments in 'geom' 

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = waikiki)

```

## Patchwork

Remember faceting? 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha =.5) +
  facet_grid(. ~ name)
```

What happens when you want multipanel plots but cant facet? 

```{r}
tmax_tmin_p =
  weather_df %>% 
    ggplot(aes(x = tmin, y = tmax, color = name)) +
    geom_point(alpha = .5) +
    theme(legend.position = "none")
  
prcp_dens_p =
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x=prcp, fill=name)) +
  geom_density(alpha=.5)

tmax_date_p = 
  weather_df %>% 
      ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none")

tmax_tmin_p + (prcp_dens_p + tmax_date_p)
```
The last line is how you join the plots so that they display together in the plots viewer and can be exported together(i think?) 


## Interplay between data manipulation and plotting
```{r}
weather_df %>% 
  mutate(
    name = factor(name),
    name = forcats::fct_relevel(name, c("Molokai_HI"))
  ) %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = .5)
```

What if I wanted densities for both tmin and tmax? 

```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperatures"
  ) %>% 
  ggplot(aes(x = temperatures, fill = observation)) + 
  geom_density(alpha = .5) +
  facet_grid(. ~ name)
```

# Revist the pups dataset

Data from the FAS study 

```{r}
pups_data = 
  read_csv("./data/FAS_pups.csv") %>% 
  janitor::clean_names() %>% 
  mutate(sex = recode(sex, `1` = "male", `2` = "female"))

litters_data = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pups_data, litters_data, by = "litter_number")
view(fas_data)

fas_data %>% 
  select(dose, day_of_tx, starts_with("pd_")) %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome",
    values_to = "pn_day"
  ) %>% 
  drop_na() %>% 
  ggplot(aes(x = dose, y = pd_ears)) +
  geom_violin() +
  facet_grid(day_of_tx ~ .)

```

